---
title: "Result Analysis for ID1 Batch 1 Full scale"
author: "Koushik Kumaraswamy"
date: "Thurday, Apr 30, 2015"
output: pdf_document
---

# Background
In this report, we  evaluate the performance of the Vector Space Model, (referred to as VSM henceforth), at full scale for ID #1, with a view of parsing the relationship betweem model accuracy and Universe representation

# Run Results and analysis
Plotting accuracy as a function of universe representation we get:

```{r runresults1,echo=FALSE}
options(warn=-1)
library(ggplot2)
fileLocation<-"C:\\Users\\koushikk\\repos\\XeevaIDM\\ID1\\ID1Batch1results.csv"
runResults<-read.csv(fileLocation)

qplot(Univrep,Accuracy,
      data=runResults,geom="smooth",
      method="loess",main="Accuracy Vs Universe Representation")

```


From the graph above, there seems to be two seperate trends we can observe.One increasing and one decreasing. We can *cut* the data set at the point of change, and observe the trends seperately as below. For the first part of the graph, we see..

```{r runresults2,echo=FALSE}
options(warn=-1)
minRows<-700 # approx region where curve reverses direction

# Process records above threshold of representation
idx<-which(runResults$Univrep>=minRows)
subsetResults<-runResults[-idx,]

qplot(Univrep,Accuracy,
      data=subsetResults,geom="smooth",
      method="loess",main="Accuracy Vs Universe Representation")

```

For the second part of the graph, we see the following...

```{r runresults3,echo=FALSE}
subsetResults<-runResults[idx,]

qplot(Univrep,Accuracy,
      data=subsetResults,geom="smooth",
      method="loess",main="Accuracy Vs Universe Representation")

```


# Analysis of data
From the above, and additional  offline analysis, we observe  that the model accuracy seems to be 95+%  if  training set count is between 700-1000. When the training set count is between 600-1200, the accuracy falls to 80+%. The accuracy further falls to 65+% if the training count varies between 500-3500. This seems to suggest that there is a sweet spot for the classifier in terms of number of records per group that it seems to do well with. This finding needs to be validated through additional randomized runs.
